### 2017-12-13 manuscriptAnalysis/MS_Analysis.Rmd
Trying to fix lack of convergence in gr model

---
title: "Growth model analysis"
output: html_notebook
---

Plots for the manuscript

```{r}
library(tidyverse)
d <- dddG[[1]]
load("D:/projects/linkedModels/data/out/ddD_bktbntats_1997.RData")
load("D:/projects/linkedModels/data/out/countsAndMasses.RData")
```

```{r custom theme}
theme_publication <- function(base_size=14, base_family="") {
  library(grid)
  library(ggthemes)
  (theme_foundation(base_size=base_size, base_family=base_family)
    + theme(plot.title = element_text(face = "bold",
                                      size = rel(1.2), hjust = 0.5),
            text = element_text(),
            panel.background = element_rect(colour = NA),
            plot.background = element_rect(colour = NA),
            axis.title = element_text(face = "bold",size = rel(1)),
            axis.title.y = element_text(angle=90,vjust =2),
            axis.title.x = element_text(vjust = -0.2),
            axis.text = element_text(),
            axis.line = element_line(colour="black"),
            axis.ticks = element_line(),
            panel.grid.major = element_line(colour="#f0f0f0"),
            panel.grid.minor = element_blank(),
            legend.key = element_rect(colour = NA),
            legend.position = "bottom",
            legend.direction = "horizontal",
            legend.key.size= unit(0.2, "cm"),
            legend.margin = unit(0, "cm"),
            legend.title = element_text(face="italic"),
            plot.margin=unit(c(10,5,5,5),"mm"),
            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
            strip.text = element_text(face="bold"),
            panel.border = element_rect(colour = "black")
    ))

}

scale_fill_Publication <- function(...){
  library(scales)
  discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

scale_colour_Publication <- function(...){
  library(scales)
  discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

```

```{r raw counts}
allFishBySpecies$riverOrderedGG <- factor(allFishBySpecies$river,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
allFishBySpecies$seasonGG <- factor(allFishBySpecies$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
allFishBySpecies$speciesGG <- factor(allFishBySpecies$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

allFish$riverOrderedGG <- factor(allFish$river,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
allFish$seasonGG <- factor(allFish$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)

ggplot(allFishBySpecies, aes(year,nAllFishBySpecies)) +
  geom_point(aes(color = speciesGG)) +
  geom_line(aes(color = speciesGG)) +
  geom_line( data = allFish, aes(year,nAllFish) ) +
  theme_publication() +
  labs( x = "Year", y = "Count" ) +
  facet_grid( seasonGG ~ riverOrderedGG)#, scales = "free_y" )

ggsave('figures/rawCount.png')
```

```{r p(detection)}
  # ggPInt <- array2df(ddD$sims.list$pBetaInt, label.x = "est")
  # 
  # ggPInt$chain <- rep(1:ddD$mcmc.info$n.chains, each = ddD$mcmc.info$n.samples/ddD$mcmc.info$n.chains)
  # ggPInt$iter <- 1:as.numeric(ddD$mcmc.info$n.samples/ddD$mcmc.info$n.chains)
  # 
  # ggplot(filter(ggPInt), aes(iter,est)) + 
  #   geom_point( aes(color=factor(chain)), size = 0.1 ) + 
  #   ylim(-1.5,1.5) + 
  #   facet_grid(d2+d4~d5+d3)

  p <- array2df(ddD$q50$pBetaInt, label.x = "est") %>%
    rename(est=est,species=d1,season=d2,river=d3,year=d4)
  ggplot(p, aes(year,invlogit(est))) +geom_point() + facet_grid(season ~ river + species)

   ggsave('figures/p.png')

  phi <- array2df(ddD$q50$phiBetaInt, label.x = "est") %>%
    rename(est=est,species=d1,season=d2,river=d3,year=d4)
  ggplot(phi, aes(year,invlogit(est))) +geom_point() + facet_grid(season ~ river + species)

# # of skunked sections    
  tmp <- dddD[[2]] %>%
           group_by(riverOrdered,year,season,section) %>%
           summarize( s = sum(enc) ) %>%
           filter( s == 0 )
  table(tmp$riverOrdered,tmp$year,tmp$season)

```

```{r overview graphs - Fig 1}
d$riverOrderedGG <- factor(d$river,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
d$seasonGG <- factor(d$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
d$speciesGG <- factor(d$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot(d, aes(year,nAllFishBySpeciesP)) +
  geom_point(aes(color=speciesGG)) +
  geom_line(aes(color=speciesGG)) +
  labs( x = "Year", y = "Estimated count") +
#  ylim(0,2500) +
  theme_publication() +
  facet_grid(seasonGG ~ riverOrderedGG, scales = "free")

ggsave('figures/count.png')

ggplot(d, aes(year,nAllFishBySpeciesP)) +
  geom_point(aes(color=species)) +
  geom_line(aes(color=species)) +
   geom_line(  aes(year,nAllFishP) ) +
  labs( x = "Year", y = "Estimated count") +
#  ylim(0,2500) +
  theme_publication() +
  facet_grid(seasonGG ~ riverOrderedGG, scales = "free")
```

```{r figs 3,4}
d %>% group_by(riverOrderedGG,seasonGG,year) %>%
  filter( riverOrderedGG == "West Brook" ) %>%
  summarize(meanF = mean(meanFlow, na.rm=T),
            sdF = sd(meanFlow, na.rm=T)) %>%
  ggplot( aes(year,meanF) ) +
  geom_point( size = 3 ) +
  geom_line() +
 # geom_smooth(method = 'lm', se=F) +
  geom_errorbar( aes(ymin = meanF - sdF, ymax = meanF + sdF, width = 0.33 ) ) +
  labs( x = "Year", y = "Stream flow (x/x)" ) +
  theme_publication() +
  facet_wrap(~seasonGG)

ggsave('figures/flow.png')

lmFunction <- function(df){ lm( meanT ~ year, data=df ) }

tempMods <- d %>%
  group_by(riverOrderedGG,seasonGG,year) %>%
  summarize(meanT = mean(meanTemperature, na.rm=T),
            sdT = sd(meanTemperature, na.rm=T)) %>%
  group_by( riverOrderedGG, seasonGG ) %>%
  nest() %>%
  mutate( model = map(data,lmFunction) )

tempModsTidy <- tempMods %>%
  mutate(tidy = map(model,broom::tidy)) %>%
  select(-data,-model) %>%
  unnest() %>%
  filter( term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*",""))

d %>% group_by(riverOrderedGG,seasonGG,year) %>%
  summarize(meanT = mean(meanTemperature, na.rm=T),
            sdT = sd(meanTemperature, na.rm=T)) %>%
  ggplot( aes(year,meanT) ) +
  geom_smooth(method = 'lm', se=F, color = 'darkgrey') +
  geom_point(  ) +
  geom_line() +
  geom_errorbar( aes(ymin = meanT - sdT, ymax = meanT + sdT, width = 0.33 ) ) +
  labs( x = "Year", y = "Stream temperature (C)" ) +
  geom_text(data=tempModsTidy,aes(x=1999,y=18, label=paste(round(estimate,2),sig))) +
  theme_publication() +
  facet_grid(seasonGG ~ riverOrderedGG)

ggsave('figures/temp.png')
```

```{r questions}

```

```{r overview graphs - Fig Appendix}

pd <- position_dodge(0.65)

# isYOY1
d %>% group_by(speciesGG,riverOrderedGG,seasonGG,year) %>%
  filter(isYOYDATA == 1) %>%
  summarize(meanGR = mean(grLength, na.rm=T),
            sdGR = sd(grLength, na.rm=T)) %>%
ggplot( aes(year,meanGR, color=speciesGG, shape=speciesGG) ) +
  geom_point( position=pd ) +
  geom_errorbar( aes(ymin = meanGR - sdGR, ymax = meanGR + sdGR), width = 0.33, position=pd ) +
  ggtitle( "isYOY = 1" ) +
  labs( x = "Year", y = "Growth rate (mm/d)" ) +
  scale_x_continuous( breaks = 1997:2015 ) +
  theme_publication() +
  theme( panel.grid.minor = element_blank(),
         axis.text.x  = element_text(angle=90, vjust=0.5, size=8) ) +
  facet_grid(seasonGG ~ riverOrderedGG)

ggsave('figures/grByYearYOY1.png')

# isYOY2
d %>% group_by(speciesGG,riverOrderedGG,seasonGG,year) %>%
  filter(isYOYDATA == 2) %>%
  summarize(meanGR = mean(grLength, na.rm=T),
            sdGR = sd(grLength, na.rm=T)) %>%
  ggplot( aes(year,meanGR, color=speciesGG, shape=speciesGG) ) +
  geom_point( position=pd ) +
  geom_errorbar( aes(ymin = meanGR - sdGR, ymax = meanGR + sdGR), width = 0.33, position=pd ) +
  ggtitle( "isYOY = 2" ) +
  labs( x = "Year", y = "Growth rate (mm/d)" ) +
  scale_x_continuous( breaks = 1997:2015 ) +
  theme_publication() +
  theme( panel.grid.minor = element_blank(),
         axis.text.x  = element_text(angle=90, vjust=0.5, size=8) ) +
  facet_grid(seasonGG ~ riverOrderedGG)

ggsave('figures/grByYearYOY2.png')

```