### 2018-04-06 manuscriptAnalysis/linearGR_ModelsMap.Rmd

---
title: "linearGR_Models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)

library(lme4)
library(tidyverse)
```

```{r load data}

load(file = paste0('D:/projects/linkedModels/data/out/dG_bktbntats1997_forLmer.RData')) # made by running mainAnalysis.R with speciesGr = c("bkt", "bnt","ats") through ddG[[ii]] <- dddG[[ii]] %>% prepareDataForJags("growth")

d2 <- data.frame(
  #sampleNumber=ddG[[1]][[2]]$sampleNumber,

  ind=ddG[[1]][[1]]$ind,
  river=ddG[[1]][[1]]$riverDATA,
  season=ddG[[1]][[1]]$season,
  year=ddG[[1]][[1]]$year,
  species=ddG[[1]][[1]]$species,
  sampleNumber=ddG[[1]][[1]]$sampleNumber,
  isYOY=ddG[[1]][[1]]$isYOYDATA,
  length=ddG[[1]][[1]]$lengthDATAStd,
  lengthByYear=ddG[[1]][[1]]$lengthDATAStd_ByYOY_Year,
  grLength=ddG[[1]][[1]]$grNotUse,
  tempStd=ddG[[1]][[1]]$tempStd,
  flowStd=ddG[[1]][[1]]$flowStd,
  countPStd=ddG[[1]][[1]]$countPAllSppStd,

  countPStdBKT=ddG[[1]][[1]]$countPStdBKT,
  countPStdBNT=ddG[[1]][[1]]$countPStdBNT,
  countPStdATS=ddG[[1]][[1]]$countPStdATS
  )

d <- d2 %>% filter(
  #  species == as.numeric(speciesInGr),
    !is.na(grLength)
  )
d <- d %>% mutate( tempStd2 = tempStd^2, flowStd2 = flowStd^2, countPStd2 = countPStd^2, length2 = length^2,
                   countPStdBKT2 = countPStdBKT^2,
                   countPStdBNT2 = countPStdBNT^2,
                   countPStdATS2 = countPStdATS^2,
                   isYOY = as.factor(isYOY), season = as.factor(season), river = as.factor(river), species = as.factor(species)
                 )
d$grLength <- ifelse(d$species == 2 & d$river == 3,NA,d$grLength)

mod <- list()

bySpecies <- d %>% 
  filter( species %in% c(1,2,3) ) %>%
  group_by(species) %>% 
  nest()

# turn factor off for ATS so the models run for all three spp
bySpecies$data[[3]]$river <- as.numeric(bySpecies$data[[3]]$river)
```

```{r}

ggplot(d, aes(sampleNumber,countPStdBKT)) + geom_point() + geom_line() + facet_grid(isYOY~river)
ggplot(d, aes(sampleNumber,countPStdBNT)) + geom_point() + geom_line() + facet_grid(isYOY~river)
ggplot(d, aes(year,countPStdBNT, color=isYOY)) + geom_point() + geom_line() + facet_grid(season~river)
ggplot(d, aes(sampleNumber,countPStdATS)) + geom_point() + geom_line() + facet_wrap(~river)

```

Try new, more flexible model structure

```{r}

mod[[1]] <- function(df) {
 lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                  isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                  isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                  isYOY * river * season * ( tempStd * flowStd * length ) +
                  isYOY * river * season * ( length * countPStdBKT ) +
                  isYOY * river * season * ( length * countPStdBNT ) +
                  isYOY * river * season * ( length * countPStdATS ) +

                  isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBKT2 ) +
                  isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBNT2 ) +
                  isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                  # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                  isYOY * river * season * ( length2 * countPStdBKT2 ) +
                  isYOY * river * season * ( length2 * countPStdBNT2 ) +
                  isYOY * river * season * ( length2 * countPStdATS2 ) +

   (1|ind), data = df ) }

mod[[2]] <- function(df) {
 lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                  isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                  isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                  # isYOY * river * season * ( tempStd * flowStd * length ) +
                  # isYOY * river * season * ( length * countPStdBKT ) +
                  # isYOY * river * season * ( length * countPStdBNT ) +
                  # isYOY * river * season * ( length * countPStdATS ) +                
                  # 
                  # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBKT2 ) +
                  # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBNT2 ) +
                  # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                  # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                  # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                  # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                  # isYOY * river * season * ( length2 * countPStdATS2 ) +

   (1|ind), data = df ) }

mod[[3]] <- function(df) {
  lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                   isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                   isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                   isYOY * river * season * ( tempStd * flowStd * length ) +
                   # isYOY * river * season * ( length * countPStdBKT ) +
                   # isYOY * river * season * ( length * countPStdBNT ) +
                   # isYOY * river * season * ( length * countPStdATS ) +

                   isYOY * river * season * ( tempStd2 * flowStd2 ) +
                   # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBNT2 ) +
                   # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                   # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                   # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                   # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                   # isYOY * river * season * ( length2 * countPStdATS2 ) +

  (1|ind), data = df ) }

mod[[4]] <- function(df) { lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                         isYOY * river * season * ( tempStd * flowStd * length ) +
                         # isYOY * river * season * ( length * countPStdBKT ) +
                         # isYOY * river * season * ( length * countPStdBNT ) +
                         # isYOY * river * season * ( length * countPStdATS ) +

                         isYOY * river * season * ( tempStd2 * flowStd2 ) +
                         isYOY * river * season * ( countPStdBKT2 ) +
                         isYOY * river * season * ( countPStdBNT2 ) +
                         isYOY * river * season * ( countPStdATS2 ) +
                         # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                         # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( length2 * countPStdATS2 ) +

                         (1|ind), data = df ) }

mod[[5]] <- function(df) { lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                         isYOY * river * season * ( tempStd * flowStd * length ) +
                         # isYOY * river * season * ( length * countPStdBKT ) +
                         # isYOY * river * season * ( length * countPStdBNT ) +
                         # isYOY * river * season * ( length * countPStdATS ) +

                         isYOY * river * season * ( tempStd2 * flowStd2 ) +
                         isYOY * river * season * ( countPStdBNT2 ) +
                         # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                         # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                         # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( length2 * countPStdATS2 ) +

                         (1|ind), data = df ) }

mod[[6]] <- function(df) { lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                         isYOY * river * season * ( tempStd * flowStd * length ) +
                         isYOY * river * season * ( length * countPStdBKT ) +
                         isYOY * river * season * ( length * countPStdBNT ) +
                         isYOY * river * season * ( length * countPStdATS ) +

                         # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                         # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                         # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( length2 * countPStdATS2 ) +

                         (1|ind), data = df ) }

mod[[7]] <- function(df) { lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                         isYOY * river * season * ( tempStd * flowStd * length ) +
                         isYOY * river * season * ( length * countPStdBKT ) +
                         isYOY * river * season * ( length * countPStdBNT ) +
                         isYOY * river * season * ( length * countPStdATS ) +

                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBKT2 ) +
                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBNT2 ) +
                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                         # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                         # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( length2 * countPStdATS2 ) +

                         (1|ind), data = df ) }

mod[[8]] <- function(df) { lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                         isYOY * river * season * ( tempStd * flowStd * length ) +
                         isYOY * river * season * ( length * countPStdBKT ) +
                         isYOY * river * season * ( length * countPStdBNT ) +
                         isYOY * river * season * ( length * countPStdATS ) +

                         isYOY * river * season * ( tempStd2 * flowStd2  ) +
                         isYOY * river * season * ( tempStd2 * flowStd2  ) +
                         isYOY * river * season * ( tempStd2 * flowStd2  ) +
                         # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                         # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( length2 * countPStdATS2 ) +

                         (1|ind), data = df ) }

mod[[9]] <- function(df) { lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                         isYOY * river * season * ( tempStd * flowStd * length ) +
                         # isYOY * river * season * ( length * countPStdBKT ) +
                         # isYOY * river * season * ( length * countPStdBNT ) +
                         # isYOY * river * season * ( length * countPStdATS ) +

                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBKT2 ) +
                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBNT2 ) +
                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                         # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                         # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( length2 * countPStdATS2 ) +

                         (1|ind), data = df ) }

mod[[10]] <- function(df) {  lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                         isYOY * river * season * ( tempStd * flowStd * length ) +
                         # isYOY * river * season * ( length * countPStdBKT ) +
                         # isYOY * river * season * ( length * countPStdBNT ) +
                         # isYOY * river * season * ( length * countPStdATS ) +

                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBKT2 ) +
                        # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBNT2 ) +
                        # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                         # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                         # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( length2 * countPStdATS2 ) +

                         (1|ind), data = df ) }

mod[[11]] <- function(df) {  lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                         isYOY * river * season * ( tempStd * flowStd * length ) +
                         # isYOY * river * season * ( length * countPStdBKT ) +
                         # isYOY * river * season * ( length * countPStdBNT ) +
                         # isYOY * river * season * ( length * countPStdATS ) +

                       #  isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBKT2 ) +
                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBNT2 ) +
                       #   isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                         # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                         # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( length2 * countPStdATS2 ) +

                         (1|ind), data = df ) }

mod[[12]] <- function(df) {  lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                         isYOY * river * season * ( tempStd * flowStd * length ) +
                         # isYOY * river * season * ( length * countPStdBKT ) +
                         # isYOY * river * season * ( length * countPStdBNT ) +
                         # isYOY * river * season * ( length * countPStdATS ) +

                        # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBKT2 ) +
                       #  isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBNT2 ) +
                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                         # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                         # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( length2 * countPStdATS2 ) +

                         (1|ind), data = df ) }

mod[[13]] <- function(df) {  lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                         isYOY * river * season * ( tempStd * flowStd * length ) +
                         # isYOY * river * season * ( length * countPStdBKT ) +
                         # isYOY * river * season * ( length * countPStdBNT ) +
                         # isYOY * river * season * ( length * countPStdATS ) +

                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBKT2 ) +
                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBNT2 ) +
                        # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                         # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                         # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( length2 * countPStdATS2 ) +

                         (1|ind), data = df ) }

mod[[14]] <- function(df) {  lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                         isYOY * river * season * ( tempStd * flowStd * length ) +
                         # isYOY * river * season * ( length * countPStdBKT ) +
                         # isYOY * river * season * ( length * countPStdBNT ) +
                         # isYOY * river * season * ( length * countPStdATS ) +

                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBKT2 ) +
                        # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBNT2 ) +
                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                         # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                         # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( length2 * countPStdATS2 ) +

                         (1|ind), data = df ) }

mod[[15]] <- function(df) {  lmer( grLength ~ isYOY * river * season * ( tempStd * flowStd * countPStdBKT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdBNT ) +
                         isYOY * river * season * ( tempStd * flowStd * countPStdATS ) +
                         isYOY * river * season * ( tempStd * flowStd * length ) +
                         # isYOY * river * season * ( length * countPStdBKT ) +
                         # isYOY * river * season * ( length * countPStdBNT ) +
                         # isYOY * river * season * ( length * countPStdATS ) +

                        # isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBKT2 ) +
                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdBNT2 ) +
                         isYOY * river * season * ( tempStd2 * flowStd2 * countPStdATS2 ) +
                         # # isYOY * river * season * ( tempStd2 * flowStd2 * length2 ) + # model error with this line in
                         # isYOY * river * season * ( length2 * countPStdBKT2 ) +
                         # isYOY * river * season * ( length2 * countPStdBNT2 ) +
                         # isYOY * river * season * ( length2 * countPStdATS2 ) +

                         (1|ind), data = df ) }

```

```{r}

bySpecies <- bySpecies %>%
   mutate(  #model1 =  map(data, mod[[1]]),
  #          model2 =  map(data, mod[[2]]),
  #          model3 =  map(data, mod[[3]]),
  #          model4 =  map(data, mod[[4]]),
  #          model5 =  map(data, mod[[5]]),
  #          model6 =  map(data, mod[[6]]),
  #          model7 =  map(data, mod[[7]]),
  #          model8 =  map(data, mod[[8]]),
          # model9 =  map(data, mod[[9]]),
          # model10 = map(data, mod[[10]]),
          # model11 = map(data, mod[[11]]),
         #  model12 = map(data, mod[[12]]),
         # model13 = map(data, mod[[13]]),
         # model14 = map(data, mod[[14]]),
          model15 = map(data, mod[[15]])
        )

o <- bySpecies
b <- AIC(o$model1[[1]],o$model2[[1]],o$model3[[1]],o$model4[[1]],o$model5[[1]],o$model6[[1]],o$model7[[1]],o$model8[[1]],o$model9[[1]],o$model10[[1]],o$model12[[1]],o$model13[[1]],o$model14[[1]]) %>% rownames_to_column() %>% arrange(AIC) #bkt
n <- AIC(o$model1[[2]],o$model2[[2]],o$model3[[2]],o$model4[[2]],o$model5[[2]],o$model6[[2]],o$model7[[2]],o$model8[[2]],o$model9[[2]],o$model10[[2]],o$model12[[2]],o$model13[[2]],o$model14[[2]]) %>% rownames_to_column() %>% arrange(AIC)#bnt
a <- AIC(o$model1[[3]],o$model2[[3]],o$model3[[3]],o$model4[[3]],o$model5[[3]],o$model6[[3]],o$model7[[3]],o$model8[[3]],o$model9[[3]],o$model10[[3]],o$model12[[3]],o$model13[[3]],o$model14[[3]]) %>% rownames_to_column() %>% arrange(AIC)#ats

(all <- data.frame(b,n,a))

```